\Large Exponential Random Graph Models with Big Networks: Maximum Pseudolikelihood Estimation and the\\Parametric Bootstrap

### 2017-04-16 TeX/paper.tex
relative running time

For both networks we first calculate the MCMLE and treat the estimated coefficients as the network's true values $\theta$. For both networks we take the same parametrization, using the number of edges, the nodal attribute for gender and the geometrically weighted edgewise shared partners (gwesp)(\cite{Hunter.2006}) distribution were we fix the decay parameter $\lambda$ at $0.5$.\\

@@

Both methods were run on a single core....\textcolor{blue}{(=> compare running time)}. The bootstrapped MPLE is not only simple and fast, it is also embarrassingly parallel, meaning that there is no problem seperating the computation into several tasks. In other words, by using multiple cores, the computing time for estimating bootstrapped MPLE confidence intervals can be reduced substantially.\\[0.3cm]

>>>>

For both networks we first calculate the MCMLE and treat the estimated coefficients as the network's true values $\theta$. For both networks we take the same parametrization, using the number of edges, the nodal attribute for gender and the geometrically weighted edgewise shared partners (gwesp)(\cite{Hunter.2006}) distribution were we fix the decay parameter $\lambda$ at $0.25$.\\

@@

The bootstrapped MPLE is not only simple and fast, it is also embarrassingly parallel, meaning that there is no problem seperating the computation into several tasks. In other words, by using multiple cores, the computing time for estimating bootstrapped MPLE confidence intervals can be reduced substantially. Figure \ref{comptime} illustrates the relative computing time of the bootstrapped MPLE using 500 networks and the MCMLE for the three networks Faux Mesa High (205 nodes), Faux Magnolia High (1461 nodes) and Supreme Court Citation (9223 nodes) for an increasing number of computing cores. For the small network we simulate $2000$ networks using a MCMC interval of $2000$ steps, for the medium network we simulate $8000$ networks using a MCMC interval of $5000$ steps and for the large network we simulate 1 million networks using $10,000$ MCMC steps in order to approximate the likelihood. The chosen sample sizes and MCMC steps are necessary to guarantee a good model fit. We define the simulation time of the bootstrapped MPLE as a function of the number of available computing cores x
$$\text{boostrapped MPLE time} = \text{network simulation time} + \cfrac{500 \cdot \text{MPLE estimation time}}{x}$$     
and define the relative computing time as
$$\text{relative computing time}= \cfrac{\text{bootstrapped MPLE time}}{\text{MCMLE time}}$$
This means that a relative computing time greater than 1 indicates that the MCMLE computing time is shorter, while a relative computing time smaller than 1 indicates that the bootstrapped MPLE provides faster results. 
\begin{figure}
\begin{center}
\includegraphics[scale=.45]{rel_time}
\caption{The y-axis gives the ratio of the bootstrapped MPLE time to that of the MCMLE time. Values below 1 indicate that the bootstrapped MPLE requires a shorter computing time.}
\label{comptime}
\end{center}
\end{figure}
Figure \ref{comptime} demonstrates that the small highschool friendship networks only require three cores for the bootstrapped MPLE to outperform the MCMLE. Even more interesting is that the even larger Supreme Court Citation network only requires two cores for the bootstrapped MPLE to outperform the MCMLE.\\[0.3cm]
